---
title: "Backtesting projections against held-out diagonals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Backtesting projections against held-out diagonals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 8,
  fig.height = 5
)
```

## Motivation

Reserving and projection methods are fitted on observed data, but their
practical value lies in how they would have performed at past valuation
dates. `backtest()` answers that question by hiding the latest `holdout`
calendar diagonals from a triangle, refitting the model on the earlier
portion, and comparing its projection to the actuals that were withheld.
This is calendar-diagonal hold-out (rather than dev-period hold-out),
because it simulates "what would the model have said *K* months ago at
the valuation date?". The cell-level metric is the Actual-Expected Gap,
$\mathrm{aeg} = v_{\mathrm{pred}} / v_{\mathrm{actual}} - 1$, where
positive values flag over-projection and negative values flag
under-projection.

## Basic usage

```{r}
library(lossratio)
data(experience)
exp     <- as_experience(experience)
tri_sur <- build_triangle(exp[cv_nm == "SUR"], cv_nm)

bt <- backtest(tri_sur, holdout = 6L, value_var = "closs", method = "mack")
print(bt)
```

The returned object is a `"backtest"` list with these key slots:

- `aeg` — per-cell `data.table` (cohort, dev, actual, pred, aeg, calendar_idx).
- `col_summary` — AEG aggregated by `dev`.
- `diag_summary` — AEG aggregated by calendar diagonal.
- `masked` — the triangle the fit was trained on (latest diagonals removed).
- `fit` — the fit object returned by `fit_fn` (a `cl_fit` or `lr_fit`).

`summary(bt)` prints the two summary tables alongside the call metadata.

## Calendar-diagonal masking limitation

Removing the latest `holdout` diagonals shortens the lower-right edge
of the triangle. A chain ladder fit on the masked triangle can only
project as far as its longest cohort × dev support; cells beyond that
support, which would belong to the very oldest cohorts at the largest
development periods, simply have no projection to compare against. The
function silently filters those unreachable cells, so `bt$aeg` always
contains only cells where both an actual and a finite projection exist.
The practical takeaway: `holdout` larger than a few diagonals reduces
the validation set fastest in the oldest cohorts, where the chain
ladder would otherwise rely on its own extrapolated tail.

## Output interpretation

**`col_summary` — systematic bias by development period.** A
consistently signed AEG at a given dev signals a structural mismatch
between the model and that maturity. Early-dev positive values usually
reflect inflated link factors; late-dev values flag tail miscalibration.

```{r}
head(bt$col_summary, 8)
```

`aeg_mean` averages cell-level AEG, `aeg_med` is the median, and
`aeg_wt = sum(pred - actual) / sum(actual)` is exposure-weighted.
Comparing the three columns flags whether a few large cells dominate
(`aeg_wt` very different from `aeg_med`) or the bias is uniform.

**`diag_summary` — calendar-year effect.** A single bad diagonal in
otherwise unbiased output points at a calendar event (a rate change,
claim handling shift, or one-off shock) that a static chain ladder
cannot see by construction.

```{r}
bt$diag_summary
```

A monotone drift across calendar diagonals (as in the SUR example
above, where AEG becomes increasingly negative across `25, ..., 30`)
typically indicates that the latest period is running better than the
earlier-cohort link factors imply.

**`aeg` — cell-level outliers.** For diagnosing specific cohort × dev
cells, inspect `bt$aeg` directly:

```{r}
head(bt$aeg, 5)
```

## Plot demos

Four plot views are registered on `"backtest"`:

```{r}
plot(bt, type = "col")    # AEG by dev (point + dashed zero line)
plot(bt, type = "diag")   # AEG by calendar diagonal
plot(bt, type = "cell")   # per-cohort AEG trajectories over dev
plot_triangle(bt)         # diverging-color heatmap on the held-out wedge
```

`type = "col"` is the right place to look for systematic dev-period
bias; `type = "diag"` reveals calendar-year drift; `type = "cell"`
exposes which cohorts contribute the bias; `plot_triangle()` puts the
cell-level AEG values on the same triangular layout as
`plot_triangle()` for the underlying fit, with a red/blue diverging
palette where red marks over-projection.

## Holdout selection

Choose `holdout` to balance two opposing effects:

- Too large: the masked triangle loses its latest experience, so the
  oldest cohorts have few or no reachable cells in their later dev
  periods. The validation set shrinks unevenly, biased toward early
  dev.
- Too small: the held-out wedge is just a thin diagonal band, which
  may not capture enough cells to reveal systematic patterns.

Typical choices are `holdout = 6L` (half-year) for monthly triangles,
or `holdout = 12L` (full year) for stronger validation when the
triangle has at least 24–30 diagonals of history.

## Choosing the fit function

`backtest()` supports both `fit_cl` and `fit_lr`. The fitter is passed
through `fit_fn`, and `value_var` selects which projection column on
`fit$full` to compare against the held-out actuals. For `fit_cl`,
`value_var` is also forwarded to the fit itself. For `fit_lr` —
which projects loss and exposure jointly — `value_var` only chooses
the comparison column, with the mapping:

| `value_var` | Compared column on `fit_lr$full` |
|-------------|----------------------------------|
| `"closs"`   | `loss_proj`                      |
| `"crp"`     | `exposure_proj`                  |
| `"clr"`     | `clr_proj`                       |

```{r}
bt_cl  <- backtest(tri_sur, holdout = 6L, fit_fn = fit_cl,
                   value_var = "closs", method = "mack")
bt_lr  <- backtest(tri_sur, holdout = 6L, fit_fn = fit_lr,
                   method = "sa", value_var = "closs")
bt_clr <- backtest(tri_sur, holdout = 6L, fit_fn = fit_lr,
                   method = "sa", value_var = "clr")

print(bt_clr)
```

Backtesting `clr` is often the more informative diagnostic: the loss
ratio is unitless and dimension-free across cohorts of very different
volume, so `aeg_mean` and `aeg_med` carry a consistent meaning across
the triangle. Backtesting `closs` weights the result toward whichever
cohorts happen to be the largest at the held-out diagonals.

## See also

- `vignette("chain-ladder")` — `fit_cl()` reference.
- `vignette("loss-ratio-methods")` — `fit_lr()` and the `"sa"`,
  `"ed"`, `"cl"` methods.
- `?backtest`, `?plot.backtest`, `?plot_triangle.backtest`.
